Trade-off Between Accuracy and Resources in CSI Feedback Compression

### 2025-04-10 Trade of between Accuracy and Resources.tex
X3-Changing paper - Structure

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{stfloats}
\usepackage{url}
\usepackage{booktabs,multirow, threeparttable}
\usepackage{multicol}
\usepackage{lipsum}% http://ctan.org/pkg/lipsum
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage{subfigure}
%\usepackage{subfig}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage{balance}
\begin{document}
\title{Robust and Efficient CSI Compression Using CLLWCsiNet in Noisy MIMO Environments}

\author{
	\IEEEauthorblockN{Fardad Ansari\IEEEauthorrefmark{1}, Dr.Mahmood Mohassel Feghhi\IEEEauthorrefmark{2}}\\
	\IEEEauthorblockA{
		\IEEEauthorrefmark{1}Faculty of Electronic and Computer Engineering, 
		\IEEEauthorrefmark{2}Faculty of Electronic and Computer Engineering, 
	}\\
	\IEEEauthorblockA{
		\IEEEauthorrefmark{1}University of Tabriz, 
		\IEEEauthorrefmark{2}University of Tabriz, 
	}\\
	\IEEEauthorblockA{
		\IEEEauthorrefmark{1}f.ansari99@ms.tabrizu.ac.ir, \IEEEauthorrefmark{2}mohasselfeghhi@tabrizu.ac.ir
	}
}

\maketitle
\begin{abstract}
Multiple-Input Multiple-Output (MIMO) systems play a crucial role in advancing 5G and 6G networks by improving spectral efficiency and data rates. In Frequency Division Duplex (FDD) systems, timely feedback of Channel State Information (CSI) from receiver to transmitter is essential but introduces significant overhead. While deep learning methods outperform traditional compressive sensing techniques, their large parameter count poses challenges for deployment on resource-limited devices, such as user equipment. Moreover, factors such as short channel coherence time and imperfect wireless channels demand more robust models.
This paper introduces CLLWCsiNet (Convolutional Latent-space Low Weight CsiNet), a novel architecture that employs convolutional latent-space learning under imperfect channel conditions. CLLWCsiNet demonstrates superior predictive accuracy in noisy environments. We provide a comprehensive overview of previous models and introduce convolutional quantization techniques, which significantly reduce parameters and enhance computational efficiency. Our approach effectively compresses CSI while maintaining predictive accuracy, making it well-suited for deployment on resource-constrained user equipment.
Simulations show that CLLWCsiNet achieves -11 dB Normalized Mean Square Error (NMSE) in low SNRs (under 20 dB) with only 70K parameters. Additionally, we evaluate the proposed model on the Quadriga dataset for the first time, and demonstrating CLLWCsiNet's robust performance across varying noise levels.

\end{abstract}

\begin{IEEEkeywords}
Massive MIMO, CSI Feedback , deep learning, Convolutional Quantization,
Dynamic Quantization.
\end{IEEEkeywords}

\section{Introduction}
Multiple-Input Multiple-Output (MIMO) systems play a pivotal role in the advancement of 5G and the forthcoming 6G networks by significantly enhancing spectral efficiency, data rates, and overall network capacity. \cite{abc} In Frequency Division Duplex (FDD) systems, accurate and timely feedback of channel state information (CSI) from the receiver to the transmitter is crucial for optimal performance. This feedback enables the transmitter to adapt its signaling strategies, thereby improving link reliability and throughput. However, the feedback process introduces substantial overhead, which can degrade system efficiency if not managed properly. Therefore, studying and mitigating feedback overhead is essential to fully realize the benefits of MIMO technology in next-generation wireless communications (Figure \ref{csifeedbackimage}) \cite{abd}.

\begin{figure}[!t]
	\centering
	\includegraphics[width=2.5in]{BTS_User_Equipement.pdf}
	\caption{Downlink From UE to Base-Station}
	\label{csifeedbackimage}
\end{figure}

CSI feedback compression is a critical component in MIMO Frequency Division Duplex (FDD) systems, particularly for enhancing spectral efficiency and network performance in 5G and emerging 6G networks. Traditional methods for CSI feedback compression, such as scalar and vector quantization \cite{vectorquantiz}, Principal Component Analysis (PCA) \cite{pca}, and codebook-based techniques \cite{compressivesensing}, have been widely utilized. These methods, while effective to some extent, often struggle with the trade-off between compression efficiency and the accuracy of CSI reconstruction. For instance, scalar and vector quantization reduce CSI dimensionality by mapping elements to predefined levels or codewords, whereas PCA retains the most significant components through orthogonal transformations. Codebook-based methods, which are integral to standards like LTE, involve selecting the closest codeword from a predefined set to represent the CSI. Despite their utility, these traditional methods may not fully capture the complex and dynamic nature of wireless channels.

Recent advancements leverage modern deep learning techniques to enhance CSI feedback compression. Methods such as autoencoders, recurrent neural networks (RNNs), and convolutional neural networks (CNNs) have demonstrated superior performance by learning efficient representations of the CSI directly from data. Autoencoders compress the CSI into a lower-dimensional latent space and reconstruct it with minimal loss, while RNNs capture temporal correlations in time-varying channels. CNNs exploit the spatial structure of the CSI, making them particularly effective for large-scale MIMO systems. More sophisticated models like Variational Autoencoders (VAEs) and Generative Adversarial Networks (GANs) provide robust compression by learning probabilistic representations, offering resilience against noise and channel imperfections. Additionally, hybrid methods combining traditional techniques with deep learning models can further enhance performance by leveraging the strengths of both approaches.

For the first time, Wen et al. \cite{abe} introduced a deep learning solution for CSI feedback compression by employing an auto-encoder architecture. Their work serves as the foundation for numerous subsequent research efforts aiming to incorporate the behavior of wireless communication systems within MIMO systems using an auto-encoder architecture. In this setup, the encoder component assumes the role of the user equipment, while the decoder portion functions as the decoder. The random spatial characteristics of channel state information have been simulated using the COST2100 channel model \cite{abf}. The COST2100 dataset has been generated accordingly. Proposed architectures have been predominantly founded upon auto-encoder structures. Wen et al. \cite{abe} introduced CsiNet, an autoencoder that employs the Residual Network \cite{abg} concept in the decoder part. Wang et al \cite{abh}. introduced time-varying CSI Feedback, aiming to simulate the coherence time in the wireless channel attributed to user equipment mobility. They incorporated Long Short-Term Memory (LSTM) units into the decoder part of an auto-encoder. Lu et al. \cite{abi} implemented the LSTM architecture in both the encoder and decoder parts of their model. They opted for a neural network instead of a convolutional form for LSTM, which resulted in a doubling of parameters and complexity compared to CsiNet-LSTM \cite{abh}, as reported. They emphasized memory for feature extraction module more than correlation by employing LSTM in parallel with a fully connected layer, akin to jump connections in residual networks \cite{abi}. Their proposed architecture, RecCsiNet \cite{abi}, achieved superior results prior to considering time correlation, by employing LSTM for residual features instead of correlation features \cite{abh}. Li and Wu \cite{abj} extended the RecCsiNet model by separating the feature extraction process from feature compression and decompression. Their proposed model, Convlstm, significantly reduces the error by a considerable amount \cite{abj}. Liu et al. \cite{abk} emphasized coherence time simulation through LSTM and proposed a new architecture called Markovenet. In addition to their suggested architecture and a simple Markov model, they introduced a new convolutional latent space that significantly reduces the number of parameters and complexity of the autoencoder. They argue that instead of using fully connected layers to extract features, which may not be crucial in the random characteristics of the channel state information matrix, a convolutional approach is more appropriate, given that only adjacent elements in the channel state information matrix exhibit strong correlations. Mashhadi et al. \cite{abl} proposed the DeepCMC network, which is a fully convolutional neural network. Dimensional compression is achieved through upsampling and downsampling in pooling layers. To further compress the representation, they utilize a context-adaptive binary arithmetic coding scheme for the encoder side of the DeepCMC output codeword. Liang et al. \cite{abm} proposed a compressive sensing architecture named CSRenet, which employs compressive sensing on the encoder side and a deep learning-based method on the decoder side. Lu et al. \cite{abn} introduced the CRNet architecture, representing one of the pioneering studies utilizing small kernel sizes for CSI matrix feature granularity extraction. They employed a residual network architecture in both the encoder and decoder. Within the residual architecture, Lu et al. utilized different kernel sizes in different paths and suggested a multi-resolution block for both encoder and decoder. An important aspect of their research is that the performance of a single-resolution refine network can be enhanced by employing different kernel sizes. Guo et al. \cite{abo} proposed a multiple-rate architecture to address the challenge posed by short coherence time intervals, where CSI must be sufficiently compressed and the auto-encoder should dynamically adjust the compression ratio according to the environment. Their proposed multiple-rate framework is capable of compressing the CSI matrix at different compression ratios without altering the architecture or requiring a new training scheme. As a result, user equipment (UE) does not need to store parameters for different architectures. Guo et al. incorporated quantization and dequantization modules in the encoder and decoder, respectively. In contrast to Lu et al. \cite{abp}, who used uniform quantization, Guo et al.\cite{abo} employed non-uniform quantization to generate binary streams at the output of the encoder. Additionally, they suggested using larger kernel sizes in both the encoder and decoder, as well as an expected residual configuration. These changes were applied to CsiNet, resulting in a new architecture called CsiNetPlus \cite{abo}. Yu et al. \cite{abq} employ the non-local block which enhance the accuracy with the help of long-range dependences. They inspired the non-local block from X et al. \cite{abr} which has been for sequence data denoising. Cai et al. \cite{abs} proposed a modified version of CsiNet called Attention-CsiNet. They replaced the fully connected layers in the encoder and decoder with LSTM modules to extract coherence among sub-carriers. Additionally, Cai et al. \cite{abs} introduced an Attention module for CSI feedback compression, inspired by the research of \cite{abt}. J et al. \cite{abt} proposed the squeeze-and-extraction network in the field of machine translation.

In this paper, we address the inefficiencies associated with the fully connected layers in many state-of-the-art models for CSI feedback compression. These layers often propagate redundant weights related to the zero parts of the channel state information, leading to a large number of parameters. This inefficiency is particularly problematic on User Equipment (UE), where resource constraints such as memory and computational power are critical factors. By focusing on quantization techniques, we aim to significantly reduce the parameter count and enhance the efficiency of CSI feedback compression models on UE.

We retrained the previously proposed models with a convolutional latent space on the COST2100 dataset \cite{abf}, revealing that many designs fail in the absence of fully connected layers. To overcome this, we proposed a new architecture named CLLWCsiNet (Convolutional Latent-space Low-weight CsiNet), which achieves high accuracy without relying on fully connected layers and performs better with significantly fewer parameters. Furthermore, to ensure generalization, we evaluated these models on the Quadriga dataset \cite{abt}, specifically generated for outdoor scenarios. Notably, many models that exhibited high accuracy on the COST2100 dataset \cite{abf} failed when tested on the Quadriga dataset\cite{abt}.

Our study also considers the imperfect channel conditions, a realistic scenario often neglected in favor of ideal conditions in wireless communication systems research. The proposed CLLWCsiNet model demonstrated performance comparable to CsiNet+DNNet \cite{abw}, which has a significantly higher number of parameters. Importantly, CLLWCsiNet maintained high accuracy across varying noise power levels, showcasing robust performance. For the first time, we employed quantization not merely for fully connected layers or bit stream generation but also within the convolutional framework. This novel application of quantization is crucial for practical deployments, as it significantly reduces the computational burden and memory requirements, making the models more suitable for resource-constrained environments typical of wireless communication systems.

The importance of quantization in this context cannot be overstated. It allows for the compression of model parameters without significant loss of accuracy, facilitating the deployment of deep learning models in real-time and low-power scenarios. Models with a high number of parameters, while potentially more accurate, often face practical limitations in wireless communication systems due to their substantial resource demands, including computational power and energy consumption. By incorporating convolutional quantization, our proposed CLLWCsiNet model addresses these challenges, offering a more efficient and scalable solution for CSI feedback compression in next-generation wireless networks.

The following is a list of the contributions:

\begin{itemize}
	\item To the best of our knowledge, this is the first time that dynamic quantization has been explicitly applied to all layers of a neural network for CSI compression tasks. Previous research has primarily focused on network binarization in fully connected layers within the latent space, aiming to enable quantization from a communication system design perspective. However, this concept can be extended to all layers, which can significantly reduce network complexity, albeit potentially at the expense of accuracy.

	Bit-level neural networks are a crucial subset of network quantization, enabling the deployment of deep learning models on devices with constrained resources, such as those found in wireless communication systems. It is important to distinguish between network quantization and codeword quantization, which involves larger compression representations as proposed in communication system designs.

	\item Unlike most research in this domain, our focus is on imperfect channels characterized by additive white Gaussian noise (AWGN) in wireless communication systems. Previous studies that consider imperfect channels include \cite{abw} and \cite{abs}. In \cite{abw}, a DNNet was proposed, which introduces high complexity even in low dimensions, despite the training process being conducted in two stages. Researchers in \cite{abs} proposed a variational auto-encoder for imperfect channels, but they did not report numerical results. In contrast, we propose a fully convolutional neural network (CNN) with a latent space alongside a simple subtraction denoising convolution filter for stabilizing accuracy and denoising, respectively. Our architecture is trained in a single stage, unlike the approach in \cite{abw}. Our results demonstrate the state-of-the-art performance of this denoising network compared to \cite{abw}.

	\item We propose a convolutional latent space low-light CSINet named CLLWCsiNet, which achieves high accuracy compared to architectures that use fully connected layers. Previous architectures rely heavily on fully connected layers, which introduce redundant zero weights and significantly increase the model's parameter count by approximately 2 million parameters. This increase in parameters leads to high computational costs Floating Point Operations (FLOPs) and storage requirements, making them impractical for user equipment with limited GPU power and storage capacity, especially in scenarios involving low temporal CSI information and CPU-based mass-produced user devices. Furthermore, most of these architectures are designed under ideal channel conditions.

\end{itemize}

\section{Modeling}

Like most scenarios adopts a single cell massive MIMO system, in which the base station equips
\begin{math}\ N_{t}\end{math} transmitting antennas as well as the UE with a single
receiving antennas. An orthogonal frequency division multiplexing (OFDM) system with \begin{math}\ N_{c}\end{math} sub-carriers for FDD downlink massive MIMO is examined. In the OFDM-MIMO system considered in this work, the subcarriers are assumed to maintain orthogonality. We will assume on  \begin{math}  {n^{th}} \end{math} sub-carriers the received signal at UE will be equal to :

\begin{equation}
	\ {y_{n}} =\textbf{h}^{H}_{n}\textbf{p}_{n} {x}_{n} + {z}_{n} 
\end{equation}
where \begin{math} {x}_{n}\in \textbf{C}\end{math},
\begin{math} {z}_{n}\in \textbf{C}\end{math},
\begin{math}\ h_{n}\in C^{N_{t}\times 1}\end{math}
and 
\begin{math}\ p_{n}\in C^{N_{t}\times 1}\end{math}
represent transmitted signal, additive Gaussian noise , channel vector and the precoding vector, respectively.
Let 
\begin{math}\ \textbf{H}\end{math}
represents the channel matrix for all sub-carriers :

\begin{equation}
		\textbf{H} =
		\left[ {\begin{array}{cccc}
				h^{H}_{1} & 0^{H}_{} & \cdots & 0^{H}_{}\\
				0^{H}_{} & h^{H}_{2} & \cdots & 0^{H}_{}\\
				\vdots & \vdots & \ddots & \vdots\\
				0^{H}_{} & 0_{H} & \cdots & h^{H}_{N_{c}}\\
		\end{array} }\right]_{N_{c}\times {N_{t}}} 
\end{equation}

For simplicity we reshape the channel matrix into
\begin{math}
	\textbf{H}= \left[ h_{1} ,h_{2},...,h_{N_{c}}\right]^{H}_{N_{t}\times N_{c}}
\end{math}
. Evidently, H has \begin{math}\ 2\times N_{c}\times N_{t}\end{math} float numbers, which is too big for direct feedback in a massive MIMO system.

\begin{math}\ \textbf{H}\end{math} is the spatial-frequency domain equivalent of the CSI. We use the 2D discrete Fourier transform (DFT) and transfer \begin{math}\ \textbf{H}\end{math} into the angle-delay domain as follows to extract the CSI features:

\begin{equation}
 \tilde{\textbf{H}}=F_{d}HF^{H}_{a}
\end{equation}

where  \begin{math} \ F_{d}\in \textbf{C}^{N_{c}\times N_{c}}\end{math} and \begin{math} \ F_{a}\in \textbf{C}^{N_{t}\times N_{t}}\end{math}.

CSI shows sparsity in the delay domain, with \begin{math} \tilde{\textbf{H}} \end{math} having significant values only in the first \begin{math} \ N_{i} \end{math}  rows because the time of arrival (TOA) between multipaths is limited in duration. Like aforementioned researches, we select  first 
\begin{math} \ N_{i} \end{math} rows of \begin{math} \tilde{\textbf{H}} \end{math} to create a new channel matrix \begin{math} \tilde{\textbf{H}_{f}} \end{math} as

\begin{equation}
% \begin{math}
 	 \tilde{\textbf{H}}_{f}=\left[ \tilde{\textbf{H}} \right]_{1:N_{i}} 
% 	 \end{math}
\end{equation}

\begin{math} \tilde{\textbf{H}}_{f} \end{math} as is still too heavy for uplink feedback, though, because in a massive MIMO system, \begin{math}\ N_{t}\end{math} is a big number. Our goal is to further compress matrix \begin{math} \tilde{\textbf{H}_{f}} \end{math} in order to minimize the weight of the feedback.

\subsection{Deep Learning Algorithm for Compression}\label{AA}

In an effort to lower transmission overheads, DL-based algorithms have recently been used in CSI feedback [11-17]. An encoder at UE first compresses the CSI data into a codeword.

\begin{equation}
 s=f_{en}(\tilde{\textbf{H}}_{f}) 
\end{equation}

and after that, a feedback channel is used to send the codeword to the BS. A decoder at the BS can recreate the CSI as :

\begin{equation}
\hat{{\tilde{\textbf{H}}}}_{f}=f_{de}(s) 
\end{equation}

The compression ration can be obtained as :

\begin{equation}
\gamma=\frac{N_{s}}{2\times N_{i}\times N_{t}}  
\end{equation}

Where \begin{math}\gamma \end{math}, \begin{math}\ N_{s} \end{math}, \begin{math}\ N_{i} \end{math} and \begin{math}\ N_{t} \end{math} are compression ratio, codeword length, selected rows from \begin{math}\tilde{H} \end{math} and number of transmit antenna.

\section{Trade of Between Accuracy and Resources} 
CNNs have demonstrated superior performance in CSI compression tasks compared to compressive sensing methods. While numerous architectures have been proposed, deploying these models on mobile hardware presents significant challenges. Consequently, it becomes imperative to strike a balance between accuracy and resource constraints when considering the practical application of CSI feedback compression on user equipment with limited resources.
Most research in CSI feedback compression focuses on various designs of auto-encoders. A significant portion of the parameters and FLOPs in auto-encoder designs is attributed to the fully connected layer. Considering the CsiNet \cite{abe} architecture, initial measurements indicate a computational load of 5.6 million FLOPs and 2.1 million trainable parameters. Upon selective removal of the fully connected layers within the architecture, these metrics were reassessed, revealing a notable reduction. Specifically, the FLOPs decreased to 3.5 million, while the number of trainable parameters plummeted to 3400.

This analysis underscores a substantial optimization opportunity through the elimination of fully connected layers. Notably, the removal of these layers resulted in a reduction of approximately 
\begin{math}37.5\%\end{math} in FLOPs, indicating a significant computational burden alleviation. Moreover, the parameter count experienced a staggering decrease of over \begin{math}99.8\%\end{math}, illuminating the dominant role played by fully connected layers in parameter-heavy neural network architectures such as CsiNet \cite{abe}.

@@

This shift underscores the profound impact of kernel size on computational demands within CNNs architectures. Notably, enlarging the kernel size to (7,7) yields a substantial rise in FLOPs, escalating by approximately 
\begin{math}362\%\end{math}. Furthermore, the adjustment marginally affects the parameter count, increasing it by a modest\begin{math}1.44\%\end{math}.

@@

CsiNet \cite{abe} achieves a parameter count under 1 million, while ACRNet \cite{abx} and CRNet \cite{abn} achieve superior Normalized Mean Square Error (NMSE) performance while maintaining a parameter count similar to CsiNet \cite{abe}. DS-NLCsiNet \cite{abq} achieves an NMSE near -17dB with nearly 10 million FLOPs and a parameter count similar to CsiNet. In contrast, Deep Decoder \cite{aby} and TransNet \cite{abz} have over 50 million FLOPs and \begin{math} 25\%\end{math} and \begin{math} 30\%\end{math} more parameters, respectively, compared to CsiNet (Figure \ref{fig:nmse-params} and \ref{fig:nmse-flops}).

@@

As highlighted by Lu et al. \cite{abp}, adjusting kernel sizes can enhance spatial correlation extraction, particularly not beneficial for CSI matrices characterized by inherent randomness due to wireless channel conditions. Furthermore, Lu et al.'s introduction\cite{abp} of network binarization represents a novel approach, marking the first application of binary neural networks \cite{aaa1} in this domain.

Many studies in this field prioritize optimizing the NMSE, overlooking the practical implications of models in real-world scenarios involving imperfect channels in wireless communication. Factors such as short-term channel correlation and resource constraints are frequently overlooked, despite their considerable dependency on model runtime. The runtime of most models is heavily reliant on the Graphics Processing Unit (GPU) of the hardware on which these models are evaluated, and this runtime is proportional to the FLOPs. Most proposed models exhibit higher NMSE when considering a snapshot of CSI and assuming zero mobility of User Equipment (UE). To address this gap, we evaluate several state-of-the-art proposed models using the QuadriGa dataset \cite{abu}. Since most models are trained on the COST2100 \cite{abf} channel model, it's essential to consider its implications for generalization to real-world scenarios. We generate CSI samples while considering the mobility of UE, and we examine snapshots at various time intervals to assess the generalization capabilities of the proposed models. We quantified the number of FLOPs and parameters required for User Equipment (UE) in Table \ref{table:methods_params}. Notably, most proposed models demand at least one million parameters, emphasizing the need to consider resource limitations and computational efficiency alongside NMSE optimization in model design. One of our proposed approaches to mitigate the high parameter count is to utilize Convolutional LatentSpace (CL), acknowledging that it may result in a slight loss of accuracy. If a significant decrease in accuracy occurs, it will demonstrate the dependence of these models on the fully connected network.

% Your content here

\begin{figure*}[!t] % Use [!t] to attempt to place the figure at the top of the page
	\centering
	\begin{minipage}[b]{0.45\textwidth} % Adjust the width as needed
		\centering
		\includegraphics[width=\textwidth]{NMSEvsNofparametters.pdf}
		\caption{NMSE vs Number of Parameters}
		\label{fig:nmse-params}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.4\textwidth} % Adjust the width as needed
		\centering
		\includegraphics[width=\textwidth]{NMSEvsFLOPs.pdf}
		\caption{NMSE vs FLOPs}
		\label{fig:nmse-flops}
	\end{minipage}
\end{figure*}
% More content here

\begin{table}[htb]
	\centering
	\caption{Methods Parameters}
	\label{table:methods_params}

	\begin{tabular}{ c|cc}
		\hline
		\multicolumn{1}{c|}{Methods}      & \multicolumn{2}{c}{Params}     \\

		& \multicolumn{1}{c}{UE Params }             & \multicolumn{1}{c}{UE FLOP} \\
		\hline
		CsiNet     & 1052K      & 1094k \\
		CRNet        & 1049K      & 1235K \\
		ACRNet-1x    & 1049K      & 1235K \\
		CsiNetPlus   & 1049K      & 1462K \\
		TransNet     & 2381K      & 1054K \\

	\end{tabular}

\end{table}

\subsection{Convolutional LatenSpace }

Liu et al. \cite{abp} employed a convolutional latent space in lieu of a fully connected layer, a departure from conventional auto-encoder architectures. Typically, fully connected neural networks constitute a significant component of such architectures, dominating in terms of both FLOPs and parameter count. Notably, the accuracy of the model is heavily influenced by the weights of these fully connected layers. However, considering that the CSI matrix exhibits significant correlation primarily with adjacent elements, a substantial portion of the parameters within the fully connected layers may be redundant, contributing to excessive FLOPs and parameter counts. This redundancy is particularly problematic in digital communication systems, which often face constraints such as time delays and storage limitations. Thus, achieving high accuracy at the expense of an inflated parameter count is impractical in such contexts.

By substituting the fully connected layer with a convolutional latent space, a significant drop in NMSE of up to 
\begin{math}50\%\end{math} was observed. We selected several of the proposed models that exhibited the highest accuracy and removed the fully connected network. Instead, we designed and trained a convolutional latent space network for them. We trained these networks alongside our proposed model from scratch with their own settings. The results are presented in Table \ref{table:convlatenspacecomp}.

As shown in Table \ref{table:convlatenspacecomp}, CLLWCsiNet achieves a performance of -15 dB, whereas CsiNet, with nearly 2 million parameters, achieves -17 dB. Without a fully connected layer, CsiNet’s \cite{abe} accuracy drops to -9.72 dB. ACRNet1x \cite{abx} (with an expansion factor of 1) also has nearly 2 million parameters and achieves -27 dB, but without a fully connected layer, its performance decreases to -8.76 dB. It is evident that the high accuracy of both ACRNet1x \cite{abx} and CRNet \cite{abn} is heavily dependent on the presence of fully connected layers.

\subsection{Proposed Model: CLLWCsiNet }
The proposed model, CLLWCsiNet, introduces a novel approach to compressing and refining CSI feedback for MIMO systems, particularly under the constraints of 5G and 6G networks. This model utilizes convolutional latent-space to reduce parameters while maintaining predictive NMSE in noisy channels, making it suitable for resource-constrained user equipment.

Model Architecture:
The CLLWCsiNet model is an extension of the CsiNet model with several key enhancements:

\subsubsection{Convolutional Latent-Space}:
This technique significantly reduces the number of parameters compared to traditional fully connected layers. This reduction is achieved by compressing the feature maps into a lower-dimensional space using convolutional layers, thus minimizing the computational overhead (Figure \ref{fig:figure1}). 
The input CSI matrix has dimensions 2 \begin{math} \times \end{math} 32 \begin{math} \times \end{math} 32, which is passed through 3 different convolutional filters with kernel sizes of (7,7), (5,5), and (3,3). Following the convolutional factorization approach proposed in \cite{abn}, the convolution kernel is split into two parts: (1,a) and (a,1), where ‘a’ is the kernel size. By factoring the convolution kernel in this manner, the model complexity is reduced while preserving the expressive power of the original convolution operation. This factorization technique has been shown to improve the efficiency of the neural network architecture without significantly sacrificing its performance.

\subsubsection{Encoder Compression}
After the initial passing of CSI matrix through the first part of the multi-resolution filtering architecture, the CSI matrix is reshaped from its original dimensions of 2 \begin{math} \times \end{math} 32 \begin{math} \times \end{math} 32 to 64 \begin{math} \times \end{math} 1 \begin{math} \times \end{math} 32. This reshaping is done through a set of selective compression filters. For a compression ratio of 8, the design selects a different set of filters, which can be switched based on the code. The compression ratio of 8 corresponds to a codeword dimension of 8 \begin{math} \times \end{math} (1 \begin{math} \times \end{math} 32). Similarly, for a compression ratio of 4, the codeword dimension becomes 16 \begin{math} \times \end{math} (1 \begin{math} \times \end{math} 32), and so on.
As shown in Figure \ref{fig:figure2}, the filtering process starts with 64 filters and progressively reduces the number of filters to 8, in two resolution passes. For a compression ratio of 4, the number of filters starts at 64 and is reduced to 16 through the multi-resolution filtering (Table \ref{table:model_comparison}).

This multi-resolution filtering approach, along with the selective compression based on the target ratio, allows the model to efficiently encode the CSI matrix while preserving the necessary information.

>>>>

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{mathtools}
\usepackage{stfloats}
\usepackage{url}
\usepackage{multirow}
\usepackage{threeparttable}
\usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{balance}

% For proper < and > symbols in text
\usepackage{textcomp}

\title{Trade-off Between Accuracy and Resources in CSI Feedback Compression}
\author{Your Name}

\begin{document}
\maketitle

\section{Trade-off Between Accuracy and Resources}

CNNs have demonstrated superior performance in CSI compression tasks compared to compressive sensing methods. While numerous architectures have been proposed, deploying these models on mobile hardware presents significant challenges. Consequently, it becomes imperative to strike a balance between accuracy and resource constraints when considering the practical application of CSI feedback compression on user equipment with limited resources.

Most research in CSI feedback compression focuses on various designs of auto-encoders. A significant portion of the parameters and FLOPs in auto-encoder designs is attributed to the fully connected layer. Considering the CsiNet \cite{abe} architecture, initial measurements indicate a computational load of 5.6 million FLOPs and 2.1 million trainable parameters. Upon selective removal of the fully connected layers within the architecture, these metrics were reassessed, revealing a notable reduction. Specifically, the FLOPs decreased to 3.5 million, while the number of trainable parameters plummeted to 3400.

This analysis underscores a substantial optimization opportunity through the elimination of fully connected layers. Notably, the removal of these layers resulted in a reduction of approximately 37.5\% in FLOPs, indicating a significant computational burden alleviation. Moreover, the parameter count experienced a staggering decrease of over 99.8\%, illuminating the dominant role played by fully connected layers in parameter-heavy neural network architectures such as CsiNet \cite{abe}.

@@

This shift underscores the profound impact of kernel size on computational demands within CNNs architectures. Notably, enlarging the kernel size to (7,7) yields a substantial rise in FLOPs, escalating by approximately 362\%. Furthermore, the adjustment marginally affects the parameter count, increasing it by a modest 1.44\%.

@@

CsiNet \cite{abe} achieves a parameter count under 1 million, while ACRNet \cite{abx} and CRNet \cite{abn} achieve superior Normalized Mean Square Error (NMSE) performance while maintaining a parameter count similar to CsiNet \cite{abe}. DS-NLCsiNet \cite{abq} achieves an NMSE near $-17$\,dB with nearly 10 million FLOPs and a parameter count similar to CsiNet. In contrast, Deep Decoder \cite{aby} and TransNet \cite{abz} have over 50 million FLOPs and 25\% and 30\% more parameters, respectively, compared to CsiNet (Figure \ref{fig:nmse-params} and \ref{fig:nmse-flops}).

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.45\textwidth]{NMSEvsNofparametters.pdf}
  \caption{NMSE vs Number of Parameters}
  \label{fig:nmse-params}
\end{figure}

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.45\textwidth]{NMSEvsFLOPs.pdf}
  \caption{NMSE vs FLOPs}
  \label{fig:nmse-flops}
\end{figure}

@@

As highlighted by Lu et al. \cite{abp}, adjusting kernel sizes can enhance spatial correlation extraction, particularly not beneficial for CSI matrices characterized by inherent randomness due to wireless channel conditions. Furthermore, Lu et al.'s introduction \cite{abp} of network binarization represents a novel approach, marking the first application of binary neural networks \cite{aaa1} in this domain.

Many studies in this field prioritize optimizing the NMSE, overlooking the practical implications of models in real-world scenarios involving imperfect channels in wireless communication. Factors such as short-term channel correlation and resource constraints are frequently overlooked, despite their considerable dependency on model runtime. The runtime of most models is heavily reliant on the Graphics Processing Unit (GPU) of the hardware on which these models are evaluated, and this runtime is proportional to the FLOPs. Most proposed models exhibit higher NMSE when considering a snapshot of CSI and assuming zero mobility of User Equipment (UE). To address this gap, we evaluate several state-of-the-art proposed models using the QuadriGa dataset \cite{abu}. Since most models are trained on the COST2100 \cite{abf} channel model, it's essential to consider its implications for generalization to real-world scenarios. We generate CSI samples while considering the mobility of UE, and we examine snapshots at various time intervals to assess the generalization capabilities of the proposed models. We quantified the number of FLOPs and parameters required for User Equipment (UE) in Table \ref{table:methods_params}. Notably, most proposed models demand at least one million parameters, emphasizing the need to consider resource limitations and computational efficiency alongside NMSE optimization in model design. One of our proposed approaches to mitigate the high parameter count is to utilize alternative compression techniques, acknowledging that it may result in a slight loss of accuracy. If a significant decrease in accuracy occurs, it will demonstrate the dependence of these models on the fully connected network.

\begin{table}[htb]
  \centering
  \caption{Methods Parameters Comparison}
  \label{table:methods_params}
  \begin{tabular}{lcc}
    \toprule
    \textbf{Method} & \textbf{UE Params} & \textbf{UE FLOPs} \\
    \midrule
    CsiNet & 1052K & 1094K \\
    CRNet & 1049K & 1235K \\
    ACRNet-1x & 1049K & 1235K \\
    CsiNetPlus & 1049K & 1462K \\
    TransNet & 2381K & 1054K \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{thebibliography}{00}
\bibitem{abe} C. Wen et al., ``Deep Learning for Massive MIMO CSI Feedback,'' \emph{IEEE Wireless Commun. Lett.}, vol. 7, no. 5, pp. 748--751, 2018.
\bibitem{abx} Z. Lu et al., ``Binarized Aggregated Network With Quantization,'' \emph{IEEE Trans. Wireless Commun.}, vol. 21, no. 7, pp. 5514--5525, 2022.
\bibitem{abn} Z. Lu et al., ``Multi-resolution CSI Feedback With Deep Learning,'' \emph{ICC}, pp. 1--6, 2020.
\bibitem{abq} X. Yu et al., ``DS-NLCsiNet: Exploiting Non-Local Neural Networks,'' \emph{IEEE Commun. Lett.}, vol. 24, no. 12, pp. 2790--2794, 2020.
\bibitem{aby} A. Chakma et al., ``Deep Decoder CsiNet for FDD Massive MIMO,'' \emph{IEEE Wireless Commun. Lett.}, vol. 12, no. 12, pp. 2073--2077, 2023.
\bibitem{abz} Y. Cui et al., ``TransNet: Full Attention Network for CSI Feedback,'' \emph{IEEE Wireless Commun. Lett.}, vol. 11, no. 5, pp. 903--907, 2022.
\bibitem{abp} Z. Lu et al., ``Binary Neural Network Aided CSI Feedback,'' \emph{IEEE Wireless Commun. Lett.}, 2020.
\bibitem{aaa1} M. Nagel et al., ``A White Paper on Neural Network Quantization,'' \emph{arXiv:2106.08295}, 2021.
\bibitem{abu} F. Burkhardt et al., ``QuaDRiGa: A MIMO Channel Model,'' \emph{EuCAP}, pp. 1274--1278, 2014.
\bibitem{abf} L. Liu et al., ``The COST 2100 MIMO Channel Model,'' \emph{IEEE Wireless Commun.}, vol. 19, pp. 92--99, 2012.
\end{thebibliography}
